---
title: "R Notebook"
---
# Data Cleaning
## Process raw data
```{r}
library(dplyr)

# Read the raw dataset
hospital <- read.csv("../data/Hospital General Information.csv")
hospital$Phone.Number <- as.character(hospital$Phone.Number)

Mortality <- hospital$Mortality.national.comparison
Safety <- hospital$Safety.of.care.national.comparison
Readmission <- hospital$Readmission.national.comparison
PatientExp <- hospital$Patient.experience.national.comparison
Effectiveness <- hospital$Effectiveness.of.care.national.comparison
Timeliness <- hospital$Timeliness.of.care.national.comparison
MedicalImagingEff <- hospital$Efficient.use.of.medical.imaging.national.comparison

# Turn "Not Available" into NA, "Above Average" into 3, "Average" into 2 and "Below Average" into 1.
levels(hospital$Efficient.use.of.medical.imaging.national.comparison) <- c(3, 1, NA, 2)
levels(hospital$Timeliness.of.care.national.comparison) <- c(3, 1, NA, 2)
levels(hospital$Effectiveness.of.care.national.comparison) <- c(3, 1, NA, 2)
levels(hospital$Patient.experience.national.comparison) <- c(3, 1, NA, 2)
levels(hospital$Readmission.national.comparison) <- c(3, 1, NA, 2)
levels(hospital$Safety.of.care.national.comparison) <- c(3, 1, NA, 2)
levels(hospital$Mortality.national.comparison) <- c(3, 1, NA, 2)
levels(hospital$Hospital.overall.rating) <- c(1, 2, 3, 4, 5, NA)
```

## Add payment column
```{r}
payment <- read.csv("../data/Payment.csv")

judge <- payment[,11]

levels(judge) <- c(3,1,2,0,0)

my.fun <- function(vec){
  vec <- as.numeric(vec)
  if(all(vec==c(0,0,0,0))){
    return(NA)
  }
  else{
    n <- 4 - sum(vec==0)
    return(sum(vec)/n)
  }
}
  
mat <- matrix(judge,ncol = 4,byrow = TRUE)
score <- apply(mat, 1, my.fun)

hospital$payment <- score
```

## Add latitude/longitude and other columns
```{r}
# Read in latitude and longitude data processed by Python codes
latlng <- read.csv("../output/laglng_transform.csv")
lat <- latlng[,1]
log <- latlng[,2]

# Bind some columns and factor the type of hospital for subsequent map marker use
hospital <- cbind(hospital, Mortality, Safety, Readmission, PatientExp, Effectiveness, Timeliness, MedicalImagingEff, lat, log)
hospital <- mutate(hospital, TF = factor(hospital$Hospital.Type, labels = c("emergency","children","critical")))

save(hospital, file="../output/hospital.RData")
```

## Delete latitude/longitude NA rows as it cannot be visualized on the map
```{r}
hos <- hospital[is.na(hospital$lat) == F & is.na(hospital$log) == F, ]
save(hos, file="../output/hos.RData")
```

## Following chunks that are greyed are our preliminary trial on transforming addresses to longtitude and latitude. We found that there are a considerable amount of NAs, therefore we write some python codes to improve this transformation process.
```{geo_codes, message=F}
# Transfer address into longtitude and latitude
# First 2500 data

library(ggmap)
library(dplyr)
hospital.geo=
  hospital[1:2500, ]%>%
  select(Address, City, State)%>%
  mutate(ADDRESS_Ext=paste(Address, City, State, sep=","))%>%
  mutate_geocode(ADDRESS_Ext)

hospital.geo$Name <- hospital$Hospital.Name[1:2500]
# Omit NA

hospital.geo <- na.omit(hospital.geo)
# Save first 2500 data
save(hospital.geo, file="~/Desktop/ads_proj2/Project2/Output/hospital_geo.RData")
```

```{geo_codes, message=F}
# Transfer address into longtitude and latitude
# Remaining data
hospital.geo.1=
  hospital[2501:nrow(hospital), ]%>%
  select(Hospital.Name, Address, City, State)%>%
  mutate(ADDRESS_Ext=paste(Hospital.Name, Address, City, State, sep=","))%>%
  mutate_geocode(ADDRESS_Ext)

# Omit NA
hospital.geo.1 <- na.omit(hospital.geo.1)
# Save remaining data
save(hospital.geo.1, file="~/Desktop/ads_proj2/Project2/Output/hospital_geo1.RData")
```

```{}
# Combine all data(first 2500 and last)
hospital.geo <- hospital.geo[, c(7, 1:6)]
hospital.geo <- hospital.geo[, c(2:7, 1)]
hospital.geo.1$Name <- hospital.geo.1$Hospital.Name 
hospital.geo.1 <- hospital.geo.1[, c(8, 2:7)]
hospital.geot <- rbind(hospital.geo, hospital.geo.1)

# Save all data
save(hospital.geot, file="~/Desktop/ads_proj2/Project2/Output/hospital_geot.RData")
```

```{}
# Merge all the necessary features and processed data into final dataset

# Merge hospital and hospital.geot into hos
hos <- merge(x = hospital, y = hospital.geot, by.x = c("Hospital.Name", "Address"), by.y = c("Name", "Address"), all.y = TRUE)
hos <- hos[, c(-20, -21)]   # delete reduplicated columns

# Change colnames of hos
colnames(hos) <- c("Name","Address","City","State","ZIP.Code",  
                   "County.Name","Phone.Number","Type",  
                   "Hospital.Ownership","Emergency Services",  
                   "Overall Rating","Mortality.national.comparison",  
                   "Safety.of.care.national.comparison",  
                   "Readmission.national.comparison",  
                   "Patient.experience.national.comparison",  
                   "Effectiveness.of.care.national.comparison" ,  
                   "Timeliness.of.care.national.comparison",  
                   "Efficient.use.of.medical.imaging.national.comparison",  
                   "payment","ADDRESS_Ext","lon","lat")                    

# Save the final dataset
save(hos, file="~/Desktop/ads_proj2/Project2/Output/hos.RData")
```